---
title: "Importing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Importing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(redcapimportr)
```

# Overview
This package is designed to import REDCap exports into R without some of the limitations that the current (as of v11.0.0) exports have. The limitations are minor, but frequently encountered by me in the course of my work. 

*NOTE*: There is an excellent package for directly interacting with the REDCap server. Ideally, this will be obsolete once the REDCapR is fully utilized here.

# File Inputs
REDCap allows a user to export all of the data in a project as an R object. Actually it produces two different files: a csv and an R script. The files themselves are named in a form similar to 
```
<Project>P_[DATA|R]_YYYY-MM-DD_HHMM
```

## Filenames
It is important to have both files in order to make things work in R; one needs both the metadata (r file) and data (csv). One practical issue that comes up when managing this type of data is that it is cumbersome to refer to both files when parameterizing the data load. However, it should be possible based on having one of the two filenames, to get both filenames.

To make the management of two components (how I refer to the two files) simpler, there are a series of functions available that culminate in `return_filename_pair()` which will return a pair of filenames, one for R and one for CSV, given either of the two filenames. For instance, 
```{r}
return_filename_pair("PartnershipTrackingP_R_2021-11-08_1628.r")
```


## R
The R script involves setting up variables, variable types, and other functions. It can be very
involved and kudos to the REDCap team for creating such an easy interface. There are, however, a
couple of small limitations in the import that would nice to fix. These are detailed below.

### Clearing the environment
The first action that the script takes is to clear out the environment. That is great when this is the
first activity you execute within an R session. If you happen to have other stuff, it may not be
as great. This isn't the end of the world, since you can just evaluate this in a separate environment
and return the results. However, it is (to me) a bit of unexpected behavior. 

### Labels for factors
Labels are defined on variables and these typically represent the question that was asked in the
REDCap form. This is very helpful, since the variable name itself may be somewhat terse or difficult
to interpret. However, in the process of loading data the R script will generously convert numeric
variables that are factors into factors. Think yes/no (as opposed to 0/1). This is a tremendous help
in figuring out how to handle specific variables. One thing that is missing here is that the labels
that were assigned to the numeric variables aren't copied over to the factor variables. A small thing,
but can be difficult to deal with in some cases. 


## CSV
The csv is fairly straightforward to load as it amounts to a table of data. This is loaded by the R
script, so not much inspection is needed of this file.

